Selecting and Controlling Physical Devices Through Head-Mounted Infrared Targeting

### 2013-09-03 doc/sigchi14/device.tex

\section{Hardware Device}

@@

Glass communicates a device id over the IR channel analogous to Patel's approach~\cite{patel_2-way_2003}. Target appliances have IR receivers and offer immediate visual feedback by toggling an LED whenever a valid id is received.

When users confirm that, further communication switches over to a ZigBee wireless network so that line of sight to the target is no longer needed.

In our prototype, Glass communicates over Bluetooth to an additional microcontroller PCB the user has to wear (Atmel ATMega256). This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency. We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change. In particular, one could substitute WiFi modules.

\subsection{Device Characterization}
We determined the usable range and accuracy empirically with one IR emitter and two IR receivers. The IR emitter constantly sent out an id signal. The receivers that correctly received the signal turn their LED on for 300 ms.

>>>>

\section{Hardware Device and Interaction}

@@

\subsection{Interaction Flow}
{\bf Look:} Users select a target device by looking in its general direction.
Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}. Target appliances have IR receivers and offer immediate visual feedback by toggling an LED whenever a valid id is received. They also respond by initiating an XBee radio connection and sending their id. This allows the Glass application to show which devices are currently targeted on the display.

{\bf Initiate:} Users confirm to initiate interaction by tapping on the Glass touchpad. The next section on disambiguation deals with cases in which multiple devices received valid IR signals. At this point, all further communication switches over to the XBee wireless network so that line of sight to the target is no longer needed (otherwise restricting head movement can become straining).

{\bf Control:} Glass displays a control interface for parameters of the chosen device. \bjoern{Describe the interaction scheme for navigating and adjusting parameters here}.
Control commands are sent over XBee radios.

{\bf Disengagement:} \bjoern{how do you end an interaction? Is there a timeout if the user forgets to manually back out of the control screen?}

\subsection{Disambiguation}
\bjoern{describe why disambiguation is necessary - combination of the relative inaccuracy of head orientation and the spread of our IR signal. Then describe how to overcome it.}

\subsection{Prototype Implementation}
In our prototype, Glass communicates over Bluetooth to an additional microcontroller board the user has to wear (Atmel ATMega256). This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency. We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change. In particular, one could substitute WiFi modules.

\section{Device Characterization}
We determined the usable range and accuracy empirically with one IR emitter and two IR receivers. The IR emitter constantly sent out an id signal. The receivers that correctly received the signal turn their LED on for 300 ms.

### 2013-09-03 doc/sigchi14/relatedwork.tex

\section{Related Work}
Relevant prior work exists in the areas of remote control of physical appliances and evaluations of pointing in physical space. We discuss each in turn.

\subsection{Remote Control of Physical Appliances}
A number of researchers propose to use different handheld devices to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or exchange other information with smart infrastructure like reading sensor netowrk data~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. Key design decisions are the method by which a target device is selected; and the method by which it is then later controlled or configured.

Several techniques use laser pointers to directly point at objects of interest. This has the advantage of also providing immediate visual feedback to the user in the shape of the laser dot. Beigl's early AIDA handheld combines laser pointing for selection with IR communication to exchange commands~\cite{beigl_point_1999}. Patel extends this technique by modulating the laser light to communicate the controllers identity~\cite{patel_2-way_2003}. Kemp et al. use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.

@@

Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in PiControl~\cite{schmidt_picontrol:_2012}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Tags like QR codes can also be affixed to devices and read by cameras. Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{lowcostsensing-uist13}. 
%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

Standard infrared remote controls for televisions and AV equipment tend to use wide-angle LEDs - since they are only meant to control a single device, there is no ambiguity  and it is unnecessary to have the user point precisely.

Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device. Selection techniques with very small selectors such as laser pointing are less appropriate for head-mounted applications. Head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as the eyes can be positioned relative to the head using the extraocular muscles. We therefore select an infrared LED - a source with a wider angle of illumination, but restrict its angle to be narrower than in general purpose IR applications.

\bjoern{Other citations to cover here: Raskar et al's radio frequency identity and geometry~\cite{raskar_rfig_2004}.}

\subsection{Evaluation of room-scale selection}
Pausch et al.'s early investigation of head-mounted displays compared orientation tracking of the head to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}. They found a clear performance benefit for head-mounted tracking.

On the other hand, Card et al. experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme. However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques. Compared to screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower.

Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using laser pointers. Various other techniques outperformed laser pointers.

Rukzio found that users strongly preferred either touching a mobile device to a target appliance to initiate interacting or pointing at a distance to ``scanning'' (initiating a Bluetooth scan and then selecting a target appliance from a list of scan results on the phone )~\cite{rukzio_experimental_2006}.

\subsection{Augmented Reality Interfaces}
\bjoern{most of the prior headmounted work is in the AR space. I don't know much about it yet.}

>>>>

\section{Related Work}
Relevant prior work exists in the areas of remote control of physical appliances, evaluations of pointing in physical space and augmented reality applications. We discuss each in turn.

\subsection{Remote Control of Physical Appliances}
Universal device control applications for smart phones are commercially available (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}), though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead. Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.

Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. Key design decisions are the method by which a target device is selected; and the method by which it is then later controlled or configured.

In several techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user what is being selected. Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}. Patel extends this technique by modulating the laser light to communicate the controllers' identity~\cite{patel_2-way_2003} to initiate radio communication. These proofs-of-concept do not include thorough evaluations. Kemp et al. use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.

@@

Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Printed tags like QR codes can also be affixed to devices and read by cameras. Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{cross_low-cost_2012}. 
%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

Standard infrared remote controls for televisions and AV equipment are only meant to control a single device and it is unnecessary to have the user point precisely. These controllers tend to use wide-angle infrared LEDs.

Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device to initiate interaction. Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications as head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes. We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.

\subsection{Evaluation of room-scale selection}
Pausch et al.'s early investigation of head-mounted displays compared head-tracking to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}. They found a clear performance benefit for head-tracking.

On the other hand, Card et al. experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme. However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques. Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.

Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using remote handheld pointing. Various techniques outperformed laser pointers.

Our work is complementary as it provides concrete performance data on using head orientation to select targets in a physical environment.

\subsection{Augmented Reality Interfaces}
Agumented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world.

%\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}
